\scuba: iterative map-making with the Sub-Millimetre User Reduction Facility

### 2013-01-07 smurf_paper.tex
address minor comments from second referee report

\begin{table}
  \caption{Summary of the model components that can be fit to \scuba\
    time-series data with SMURF. Only the first group of models are
    typically fit to the data (\model{COM}--\model{NOI}) in the
    indicated order. The remaining models (\model{DKS}--\model{PLN})
    are usually omitted, although they are available as options.}
  \vspace{0.2cm}
  \centering
  \begin{tabular}{c|l}
    \hline
    Model & Description \\
    \hline
    \model{COM} & remove common-mode signal \\
    \model{GAI} & common-mode scaled to each bolometer \\
    \model{EXT} & extinction correction \\
    \model{FLT} & Fourier transform filter \\
    \model{AST} & map estimate of astronomical signal \\
    \model{NOI} & noise estimation \\
    \hline
    \model{DKS} & dark squid cleaning along columns \\
    \model{PLN} & 2-dimensional time-varying plane removal \\
    \model{SMO} & time-domain smoothing filter \\
    \model{TMP} & pointing as baseline template \\
    \hline
    \end{tabular}
  \label{tab:components}
\end{table}

@@

\subsubsection{\model{FLT}: Fourier transform filter}
\label{sec:flt}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{lockman_psf}
\caption{Slice through the angular response to an ideal Gaussian point
  source (solid line) along the R.A. axis following map-making using
  the blank-field processing configuration with SMURF (dotted line),
  and upon application of the whitening filter (dashed line). The
  ``map filtered'' response is produced by adding the ideal Gaussian
  to the real data near the centre of the Lockman Hole map, and
  re-reducing the data. The resulting dotted line gives the expected
  shape of a source in panel (a) from
  Fig.~\ref{fig:lockman_maps}. Applying the whitening filter
  (reciprocal of the solid orange line in
  Fig.~\ref{fig:lockman_pspec}) to the dotted line gives the whitened
  line (dashed), which is the expected shape of a source in panel (b)
  from Fig.~\ref{fig:lockman_maps}. Finally, cross-correlating the
  whitened map with this whitened PSF is an effective ``matched
  filter'' for identifying point-like sources, and this smoothed map
  is shown in panel (c) of Fig.~\ref{fig:lockman_maps}.}
\label{fig:lockman_psf}
\end{figure}

@@

To suppress noise in the map, we construct a whitening filter from the
reciprocal of the jackknife angular power spectrum (orange line in
Fig.~\ref{fig:lockman_pspec}), normalised by the white-noise level
estimated from the RMS power at angular frequencies $>
0.1$\,arcsec$^{-1}$. The filter is applied by scaling the power
spectrum of the map by this function, and then transforming back into
real space. The whitened map is shown in Fig.~\ref{fig:lockman_maps}b,
and for comparison, the jackknife map has also been whitened in
Fig.~\ref{fig:lockman_maps}e. In both cases, the maps are visibly
flatter than the non-whitened cases.

@@

For blind, high-redshift surveys, individual sources are expected to
be un-resolved by the \scuba\ 7.5--14.5\,arcsec FWHM beams. Under this
assumption, and also assuming that the map noise is white,
cross-correlation between the map and the known PSF, or matched
filtering, yields the maximum-likelihood flux density of supposed
point-sources centred over every location in the resulting map
\citep[an extremely well-known result throughout astronomy,
see][]{stetson1987}. Peak identification in such smoothed maps have
been used extensively in the submillimetre community, as both an
efficient source-detection and photometric measurement strategy. In
terms of the angular power spectrum, a matched filter may also be
thought of as an optimal low-pass filter, that suppresses noise on
scales smaller (and therefore at higher angular frequencies) than the
beam. For the case at hand, we may use this cross-correlation
technique since the map has been whitened. However, we must first
establish the effective shape of point sources in this map due both to
map-making itself, and the whitening filter. We determine the effect
of map-making by adding a synthetic (and high-\snr) point source to
the real data, and measure its resulting shape in the map (solid and
dotted lines in Fig.~\ref{fig:lockman_psf}, respectively). Next, the
Fourier transform of the map filtered PSF is divided by the jackknife
noise power spectrum to calculate the final whitened PSF (dashed line
in Fig.~\ref{fig:lockman_psf}). Maps smoothed by this shape are shown
for the whitened signal, and jackknife maps in
Figs.~\ref{fig:lockman_maps}c and f, respectively. Note that these
images are plotted in \snr\ units, where the smooth noise maps have
been calculated by propagating the original noise maps output by SMURF
both through the whitening and matched filters (each of which are
linear operations).

>>>>

\begin{table}
  \caption{Summary of the model components that can be fit to \scuba\
    time-series data with SMURF. Only the first group of models are
    typically fit to the data (\model{COM}--\model{NOI}) in the
    indicated order. The remaining models (\model{DKS}--\model{PLN})
    are usually omitted, although they are available as options.}
  \vspace{0.2cm}
  \centering
  \begin{tabular}{c|l}
    \hline
    Model & Description \\
    \hline
    \model{COM} & remove common-mode signal \\
    \model{GAI} & common-mode scaled to each bolometer \\
    \model{EXT} & extinction correction \\
    \model{FLT} & Fourier Transform filter \\
    \model{AST} & map estimate of astronomical signal \\
    \model{NOI} & noise estimation \\
    \hline
    \model{DKS} & dark squid cleaning along columns \\
    \model{PLN} & 2-dimensional time-varying plane removal \\
    \model{SMO} & time-domain smoothing filter \\
    \model{TMP} & pointing as baseline template \\
    \hline
    \end{tabular}
  \label{tab:components}
\end{table}

@@

\subsubsection{\model{FLT}: Fourier Transform filter}
\label{sec:flt}

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{lockman_psf}
\caption{Slice through the angular response to an ideal Gaussian point
  source (solid line) along the R.A. axis following map-making using
  the blank-field processing configuration with SMURF (dotted line),
  and upon application of the whitening filter (dashed line). The
  ``map filtered'' response is produced by adding the ideal Gaussian
  to the real data near the centre of the Lockman Hole map, and
  re-reducing the data. The resulting dotted line gives the expected
  shape of a source in panel (a) from
  Fig.~\ref{fig:lockman_maps}. Applying the whitening filter to the
  dotted line (dividing the Fourier Transform of the map filtered
  response by the square root of the solid orange line in
  Fig.~\ref{fig:lockman_pspec}, and then transforming back to real
  space) gives the whitened line (dashed), which is the expected shape
  of a source in panel (b) from Fig.~\ref{fig:lockman_maps}. Finally,
  cross-correlating the whitened map with this whitened PSF is an
  effective ``matched filter'' for identifying point-like sources, and
  this smoothed map is shown in panel (c) of
  Fig.~\ref{fig:lockman_maps}.}
\label{fig:lockman_psf}
\end{figure}

@@

To suppress noise in the map, we construct a whitening filter whereby
the Fourier Transform of the map is divided by the square root of the
jackknife power spectrum (orange line in
Fig.~\ref{fig:lockman_pspec}), normalised by the white-noise level
estimated from the RMS power at angular frequencies $>
0.1$\,arcsec$^{-1}$, and transforming back into real space. The
whitened map is shown in Fig.~\ref{fig:lockman_maps}b, and for
comparison, the jackknife map has also been whitened in
Fig.~\ref{fig:lockman_maps}e. In both cases, the maps are visibly
flatter than the non-whitened cases.

@@

For blind, high-redshift surveys, individual sources are expected to
be un-resolved by the \scuba\ 7.5--14.5\,arcsec FWHM beams. Under this
assumption, and also assuming that the map noise is white,
cross-correlation between the map and the known PSF, or matched
filtering, yields the maximum-likelihood flux density of supposed
point-sources centred over every location in the resulting map
\citep[an extremely well-known result throughout astronomy,
see][]{stetson1987}. Peak identification in such smoothed maps have
been used extensively in the submillimetre community, as both an
efficient source-detection and photometric measurement strategy.  For
the case at hand, we may use this cross-correlation technique since
the map has been whitened. However, we must first establish the
effective shape of point sources in this map due both to map-making
itself, and the whitening filter. We determine the effect of
map-making by adding a synthetic (and high-\snr) point source to the
real data, and measure its resulting shape in the map (solid and
dotted lines in Fig.~\ref{fig:lockman_psf}, respectively). Next, the
Fourier Transform of the map filtered PSF is divided by the square
root of the jackknife noise power spectrum to calculate the final
whitened PSF (dashed line in Fig.~\ref{fig:lockman_psf}). Maps
smoothed by this shape are shown for the whitened signal, and
jackknife maps in Figs.~\ref{fig:lockman_maps}c and f,
respectively. Note that these images are plotted in \snr\ units, where
the smooth noise maps have been calculated by propagating the original
noise maps output by SMURF both through the whitening and matched
filters (each of which are linear operations). In terms of the angular
power spectra, this complete process can be thought of as an optimal
band-pass filter that has both suppressed low-frequency noise, and
information on scales that are smaller (higher frequencies) than the
beam.
